---
title: "Getting Started with dineExpand"
author: "Gopi Nath Vajpai"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with dineExpand}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

## Why dineExpand?

Dictionary-based text analysis is widely used in hospitality research —
but building a dictionary that captures how diners *actually* write is
difficult. General-purpose embeddings trained on Wikipedia or news text
don't know that "umami" belongs near "flavor", that "gastropub" relates
to "ambience", or that "waitstaff" is central to service quality.

`dineExpand` solves this with word embeddings trained on **512,140 Yelp
restaurant reviews**. Any hospitality construct — revisit intention, food
neophobia, authenticity, price perception — can be expanded with a single
function call.

---

## Installation

```{r install}
# install.packages("remotes")
remotes::install_github("gvajpai/dineExpand")
```

---

## Step 1 — Load restaurant vectors (once)

The vectors are downloaded from GitHub (~33 MB) and cached permanently.
All future calls load from cache in seconds.

```{r load}
library(dineExpand)

vecs <- load_dining_vectors()
#> ! Downloading restaurant GloVe vectors (~33 MB). This happens only once.
#> i Trained on 512,140 Yelp restaurant reviews.
#> v Vectors cached at: ~/.local/share/dineExpand/yelp_restaurant_glove_100d.rds
#>   Loaded 98 432 words × 100 dimensions.
```

---

## Example constructs

### Revisit intention

```{r revisit}
expand_dining(
  seed        = c("revisit", "return", "loyalty"),
  vectors     = vecs,
  interactive = FALSE
)
#>          word similarity      seed pct_match
#> 1      repeat      0.831    return       83%
#> 2  definitely      0.819    return       82%
#> 3  allegiance      0.812    loyalty      81%
#> 4      coming      0.803    return       80%
#> 5  patronize   0.797    revisit      80%
```

### Food neophobia

```{r neophobia}
expand_dining(
  seed        = c("adventurous", "novel", "unfamiliar", "exotic"),
  seed_mode   = "centroid",
  vectors     = vecs,
  interactive = FALSE
)
```

### Authenticity

```{r authenticity}
expand_dining(
  seed        = c("authentic", "traditional", "genuine", "local"),
  vectors     = vecs,
  interactive = FALSE
)
```

### Service quality

```{r service}
expand_dining(
  seed        = c("attentive", "courteous", "prompt", "helpful"),
  vectors     = vecs,
  interactive = FALSE
)
```

---

## Expanding a full lexicon

Supply any two-column data frame (word + dimension) to expand all
dimensions at once:

```{r lexicon}
my_dict <- data.frame(
  word      = c("aroma",     "flavor",    "texture",
                "delighted", "excited",   "moved",
                "revisit",   "return",    "loyal"),
  dimension = c("sensory",   "sensory",   "sensory",
                "affect",    "affect",    "affect",
                "revisit",   "revisit",   "revisit")
)

candidates <- expand_dining_lexicon(my_dict, vectors = vecs, threshold = 0.65)
table(candidates$dimension)
```

---

## Export and score

```{r export}
library(quanteda)

# Export to quanteda dictionary
dict <- lexiExpand::export_dict(
  candidates,
  seed   = my_dict$word,
  name   = "dining_constructs",
  format = "quanteda"
)

# Score reviews
reviews <- corpus(c(
  r1 = "The aroma was captivating and the flavors extraordinary.",
  r2 = "We will definitely return — truly loyal customers now.",
  r3 = "Staff were attentive, courteous, and incredibly prompt."
))

dfm(tokens(reviews)) |>
  dfm_lookup(dictionary = dict) |>
  convert(to = "data.frame")
```

---

## Train on your own corpus

```{r train}
library(data.table)
reviews <- fread("my_restaurant_reviews.csv")

vecs_custom <- train_dining_vectors(
  texts      = reviews$text,
  dims       = 100L,
  cache_path = "~/my_vectors.rds"
)

expand_dining(c("aroma", "flavor"), vectors = vecs_custom)
```
